---
title: "Final Project"
author: 'Nathanael Nam SID: 005751279 Lecture: 1 Section: 1A'
date: '2022-07-20'
output: pdf_document
---

```{r}
cars_test <- read.csv("SummercarsTestNoY.csv")
cars_train <- read.csv("SummercarsTrain.csv")
library(corrplot)
library(ggplot2)
library(ggfortify)
library(GGally)
library(car)
library(leaps)
library(stringr)
library(MASS)
```

Overview of the Data


```{r}
dim(cars_train)

names(cars_train)

summary(cars_train$PriceNew)
ggplot(data = cars_train, aes(y = PriceNew)) + geom_boxplot() + ggtitle("Boxplot of PriceNew Variable")
```




```{r}
corr_list_numeric <- round(cor(x = cars_train[sapply(cars_train, is.numeric)], y = cars_train$PriceNew), 3)
corr_list_numeric

full.mod <- lm(PriceNew ~., data = cars_train)
summary(full.mod)
autoplot(full.mod)
```

Creating an initial model based off of high R Squared for the numeric variables using the leaps package

```{r}
l <- cars_train[, sapply(cars_train, is.numeric)]
t <- l[, 1:15]
out <- regsubsets(x = t, y = cars_train$PriceNew, method = "exhaustive", nvmax = 16)
summary(out)


plot(out, scale = "adjr2")
plot(out, scale = "bic")

plot(1:15,summary(out)$bic) 
lines(1:15,summary(out)$bic) 


plot(1:15,summary(out)$adjr2) 
lines(1:15,summary(out)$adjr2) 

which.max(summary(out)$adjr2)
which.min(summary(out)$bic)
which.min(summary(out)$cp)

out.f <- regsubsets(x = t, y = cars_train$PriceNew, method = "forward", nvmax = 16)
summary(out.f)


plot(out.f, scale = "bic")
plot(out.f, scale = "adjr2")

which.max(summary(out.f)$adjr2)
which.min(summary(out.f)$bic)
which.min(summary(out.f)$cp)

out.b <- regsubsets(x = t, y = cars_train$PriceNew, method = "backward", nvmax = 16)
summary(out.b)
plot(out.b, scale = "bic")
plot(out.b, scale = "adjr2")

which.max(summary(out.b)$adjr2)
which.min(summary(out.b)$bic)
which.min(summary(out.b)$cp)
```

Through observation we can see that our best variables are MPG.highway, Horsepower, RPM, Rev.per.mile, Passengers, Length, Width, Turn.circle, Wheelbase


We will create an initial linear model with these variables

```{r}
m1 <- lm(PriceNew ~ MPG.highway + Horsepower + RPM + Rev.per.mile + Passengers + Length + Width + Turn.circle, data = cars_train)
summary(m1)
```


Add the categorical variables with the highest correlation to PriceNew
The highest correlation of the categorical variables are Make, Type, Origin, AirBags
Since we cannot use Make by itself I have modidifed it to depict the region in which the car was manufactured

```{r}
cars_train$MakeNew <- character(length(cars_train$Make))
for (i in seq_len(length(cars_train$MakeNew))){
if(word(cars_train$Make[i], 1) %in% c("Acura", "Honda", "Lexus", "Infiniti", "Mazda", "Mitsubishi", "Subaru", "Toyota", "Nissan", "Suzuki", "Hyundai", "Kia")) {cars_train$MakeNew[i] <- "Asia"}

if(word(cars_train$Make[i], 1) %in% c("Audi", "BMW", "Mercedes-Benz", "Volkswagen", "Saab", "Volvo")) {cars_train$MakeNew[i] <- "Europe"}

if(word(cars_train$Make[i], 1) %in% c("Buick","Cadillac", "Chevrolet", "Chrysler", "Dodge", "Eagle", "Ford", "Geo", "Lincoln", "Mercury", "Oldsmobile", "Plymouth", "Pontiac", "Saturn")) {cars_train$MakeNew[i] <- "USA"}
}


cars_test$MakeNew <- character(length(cars_test$Make))
for (i in seq_len(length(cars_test$MakeNew))){
if(word(cars_test$Make[i], 1) %in% c("Acura", "Honda", "Lexus", "Infiniti", "Mazda", "Mitsubishi", "Subaru", "Toyota", "Nissan", "Suzuki", "Hyundai", "Kia")) {cars_test$MakeNew[i] <- "Asia"}

if(word(cars_test$Make[i], 1) %in% c("Audi", "BMW", "Mercedes-Benz", "Volkswagen", "Saab", "Volvo")) {cars_test$MakeNew[i] <- "Europe"}

if(word(cars_test$Make[i], 1) %in% c("Buick","Cadillac", "Chevrolet", "Chrysler", "Dodge", "Eagle", "Ford", "Geo", "Lincoln", "Mercury", "Oldsmobile", "Plymouth", "Pontiac", "Saturn")) {cars_test$MakeNew[i] <- "USA"}
}
```

Create a new category in the dataset with the region of origin based on the brand of the car.

Testing the new linear regression models and checking for conditions.

```{r}
m2 <- lm(PriceNew ~ MPG.highway + Horsepower + RPM + Rev.per.mile + Passengers + Length + Width + Turn.circle + MakeNew + Type + AirBags, data = cars_train)
summary(m2)
vif(m2) 
autoplot(m2)
```


```{r}
m3 <- lm(PriceNew ~ MPG.highway + Horsepower + RPM + Rev.per.mile + Length + MakeNew + Type + AirBags, data = cars_train)
summary(m3)
vif(m3)
```


```{r}
m4 <- lm(PriceNew ~ MPG.highway + Horsepower + Rev.per.mile + Length + + MakeNew + AirBags, data = cars_train)
summary(m4)
vif(m4)

```

Removing and testing variables to find a better model:


Now we can start transforming variables because of the conditions.

```{r}
summary(powerTransform(cbind(PriceNew,Horsepower,MPG.highway,Length)~1,data=cars_train))

m5 <- lm(PriceNew^(-.5) ~ Horsepower + Length + MakeNew + AirBags + Type, data = cars_train)
summary(m5)
vif(m5)
autoplot(m5)
```


```{r}
autoplot(m5)[[1]]
autoplot(m5)[[2]]
autoplot(m5)[[3]]
autoplot(m5)[[4]]
leveragePlots(m5)

```



```{r}
cook<-cooks.distance(m5) 
qplot(fitted.values(m5),cook)+geom_hline(yintercept =  4/(length(cars_train$PriceNew)-4),col="red", linetype="dashed") + ggtitle("Cook's Distance against Fitted Values") + scale_x_continuous(name = "Fitted Values") + scale_y_continuous(name = "Cook's Distance")



STD.Res<-rstandard(m5) 
qplot(fitted.values(m5),STD.Res)+geom_hline(yintercept =c(-2,2),col="red", linetype="dashed") + ggtitle("STDRes against Fitted Values") + scale_x_continuous(name = "Fitted Values") + scale_y_continuous(name = "STDRes")


cooksd <- cooks.distance(m5)
sample_size <- nrow(cars_train)
influential <- as.numeric(names(cooksd)[(cooksd > (4/sample_size))])

cars_train_new <- cars_train[-influential, ]
```

```{r}
m6 <- lm(PriceNew^(-.5) ~ Horsepower + Length + MakeNew + AirBags + Type, data = cars_train_new)
summary(m6)
vif(m6)
leveragePlots(m6)
autoplot(m6)


ggpairs(cars_train[c(3,8,14,23)])

attempt_3 <- (predict(m6, newdata = cars_test))^(-2)
summary(attempt_3)

write.csv(attempt_3, "/Users/nathanaelnam/Desktop/Stats 101A/Week 4/Attempt 3.csv")


m7 <- lm(PriceNew ~ Horsepower + Length + MakeNew + AirBags + Type, data = cars_train)
autoplot(m7)
autoplot(m7)[[1]]
autoplot(m7)[[2]]
autoplot(m7)[[3]]
autoplot(m7)[[4]]
vif(m7)
leveragePlots(m7)

```


```{r}


write.csv(cars_train_new, "/Users/nathanaelnam/Desktop/Stats 101A/cars_train_new.csv")

cars_test_new <- cars_test
write.csv(cars_test_new, "/Users/nathanaelnam/Desktop/Stats 101A/cars_test_new.csv")
```


